---
title: "Package gssquery"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(gssquery)
```
1. What the package does?

This package can be used to perform fixed effect model where the treatment status is an indicator with two levels, and outcome variable is numeric. Compared to random effect model, a fixed effect model allows one to control for time-invariant unobserved individual characteristics that can be correlated with the observed independent variables. This model is commonly used in Economics especially in the field of health economics. 

Further analyses can be also performed to check the robustness of the main effect. Specifically, event study is included to check if there is a lag or lead of the main effect. In addition, falsification exercises (1) leave one out is included to test the sensitivity of the main effect to subset of sample; (2) random treatment status reshuffling is included to test the sensitivity of the main effect to treatment status reshuffling.

Non-quantitative functions such as get attributes, query attributes, plot attribute of a given data is also included.

Deviation from proposal: The proposed design of this package is for GSS specifically. The final package can be used for any data with similar structure.


2. Use cases.

Gerneral Social Survey (GSS) data from `gss.norc.org` is included in this package to test its applicability. The raw data is pre-processed as in "/data-raw/GSS.R" to fit the package use, as the raw data is annually, not panel. 

The package can be used to test if US presidential election affects household happiness. The corresponding data for this case is included in `GSS.rda`.

The corresponding arguments using data GSS are:

y: outcome variable, which is `happy.score` evaluating how happy a household is.
x: treatment variable, which is `election`, or `re.election`.
FE: fixed effects, which is `year`.
time: time fixed effect, which is `year`.
out: group variable indicating subgroup to be dropped, which can be any attribute.


2.1. The data `GSS` contains hundreds of attributes. Function `GerAttributes` can be used to show attribute names, so that the user can decide which attribute is useful. 

```{r}
data(GSS)
GetAttributes(data = GSS)
```
For the above effect, related variables are outcome variable `happy.score`, treatment variable `election` which takes value for election years and 0 otherwise, and `re.election` which takes value 2 for re-election years，1 for other election years and 0 otherwise, time fixed effect `year`, and household characteristics affecting election, such as `age`, `income`, `health`, `educ`.


2.2. The function QueryAttributes can be used to query any attribute of interest. 

```{r}
query<-QueryAttributes(data = GSS,'happy','health','income')
head(query)
```

2.3. The function `GetCleanLmData` can be used to get attributes of interest, which is cleaned by replacing any value with nonsense by NA. Since `GSS` is a survey data, this is a common issue. The data has to be cleaned before running any quantitative analyses.

```{r}
LmData <- GetCleanLmData(data = GSS, y = 'happy.score', x = 'election', FE = 'year', common.na.strings = c("DK", "IAP", "UNCODEABLE & IAP", ".a", ".b","DK NA"), 'age', 'income', 'health', 'educ')
head(LmData)
LmData.reelec <- GetCleanLmData(data = GSS[GSS$re.election != 1,], y = 'happy.score', x = 're.election', FE = 'year', common.na.strings = c("DK", "IAP", "UNCODEABLE & IAP", ".a", ".b","DK NA"), 'age', 'income', 'health', 'educ')
head(LmData.reelec)
LmData.elec <- GetCleanLmData(data = GSS[GSS$re.election != 2,], y = 'happy.score', x = 're.election', FE = 'year', common.na.strings = c("DK", "IAP", "UNCODEABLE & IAP", ".a", ".b","DK NA"), 'age', 'income', 'health', 'educ')
head(LmData.elec)
```

2.4. The data returned from `GetCleanLmData` can be used for quantitative analyses.

2.4.1. The function `Analysis.lm` can be used to perform fixed effect model, with selected fixed effect, and other control variables.

```{r}
library(dplyr)
LmData <- LmData %>% mutate(age = as.numeric(age)) %>% filter(age<=60)
summary(Analysis.lm(data = LmData, y = 'happy.score', FE = 'year'))

```
The results indicate households on average are happier by 0.077 in election years (campaign for brand new presidents or re-election) compared to non-election years.

```{r}
LmData.reelec <- LmData.reelec %>% mutate(age = as.numeric(age)) %>% filter(age<=60)
summary(Analysis.lm(data = LmData.reelec, y = 'happy.score', FE = 'year'))

```
The results indicate households on average are happier by 0.55 in re-election years compared to  non-election years.

```{r}
LmData.elec <- LmData.elec %>% mutate(age = as.numeric(age)) %>% filter(age<=60)
summary(Analysis.lm(data = LmData.elec, y = 'happy.score', FE = 'year'))
```
The results indicate households on average are happier by 0.077 in elections years (campaign for brand new presidents) compared to non-election years.


2.4.2. The function `LagEffect` can be used to check if there is a lag effect.

```{r}
LagEffect(data = LmData, y = 'happy.score', x = 'election', time = 'year', FE = 'year')


```
The results indicate the election (for brand new presidents or re-election) causes households to be less happy by 1.04 two years later.

```{r}
LagEffect(data = LmData.elec, y = 'happy.score', x = 're.election', time = 'year', FE = 'year')

```
The results indicate the election (for brand new presidents) causes households to be less happy by 1.04 two years later.
```{r}
LagEffect(data = LmData.reelec, y = 'happy.score', x = 're.election', time = 'year', FE = 'year')

```
The results indicate the re-election causes households to be happier by 0.37 two years later.


2.4.3. The function `LeadEffect` can be used to check if there is lead effect.

```{r}
LeadEffect(data = LmData, y = 'happy.score', x = 'election', time = 'year', FE = 'year')

```
The results indicate household happiness in two years before the election has no statistically difference between election (for brand new president or re-election) years and non-election years .

```{r}
LeadEffect(data = LmData.elec, y = 'happy.score', x = 're.election', time = 'year', FE = 'year')
```
The results indicate household happiness in two years before the election has no statistically difference between election (for brand new president) years and non-election years .


```{r}
LeadEffect(data = LmData.reelec, y = 'happy.score', x = 're.election', time = 'year', FE = 'year')

```
The results indicate households are less happy by 0.55 two years before re-election.

2.4.4. The function `PlaceboTSReshuffle` can be used to check if the effect from original treatment status is an outlier from random reshuffling of treatment status. If the effect is truely a causal effect, then the coefficient from original treatment status should stand out.

```{r}
PlaceboTSReshuffle(data = LmData, y = 'happy.score', x = 'election', FE = 'year')
PlaceboTSReshuffle(data = LmData.elec, y = 'happy.score', x = 're.election', FE = 'year')
PlaceboTSReshuffle(data = LmData.reelec, y = 'happy.score', x = 're.election', FE = 'year')


```
The graph suggests that only the effect of re-election on household happiness passed this falsification exercise since the red dot is an outlier

2.4.5. The function `PlaceboLOU` can be used to check if the main effect is sensitive to subset of sample. The argument `out` specifies group variable. For example, `out = "year"`, then the main effect will be re-examined for sub sample where one year is droped.

```{r}
PlaceboLOU(data = LmData, y = 'happy.score', x = 'election', FE = 'year', out = 'year')
PlaceboLOU(data = LmData.elec, y = 'happy.score', x = 're.election', FE = 'year', out = 'year')
PlaceboLOU(data = LmData.reelec, y = 'happy.score', x = 're.election', FE = 'year', out = 'year')

```
The results indicate that the direction of the effect of US presidential re-election on household happiness is reversed after dropping year 2002. And the effect of US presidential election on household happiness is much larger after dropping year 2000.


2.5. The function `PlotAttribute` can be used to plot a chosen attribute with given group variable. GSS data are almost all categorical with level fewer than 15, so the function does a bar plot by year group. For a small exceptions like `age`, this function plot the mean by group.

This visualization shows the trend (sample distribution) of attribute. 

```{r}
PlotAttribute(data = GSS, attr = "happy", group.by = "year")
PlotAttribute(data = GSS, attr = "age", group.by = "year")
```
